---
title: "a3_task3_katelyn_toigo"
author: "Katelyn Toigo"
date: "10/31/2020"
output: html_document
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

## Summary

I will be analyzing the urchins.csv data.  I will be using version control, project management, data wrangling, analysis, and visualization skills.

## Read in the data

Read in the urchins.csv file and also clean the names so that they are all in lowercase snakecase.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Exploratory data visualization

### Faceted histogram

Comparing common_name vs. size
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram() +
  facet_wrap(~common_name) +
  labs(x = "Size (cm)", y = "Count")
```

### Boxplot

Common name vs. size
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot()+
  labs(x = "Common Name", y = "Size (cm)")
```

## Descriptive Statistics Table

```{r, echo=TRUE, message=FALSE, warning=FALSE}
urchin_table <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n())
  kbl(urchin_table) %>% 
  kable_styling()
```

