---
title: "a3_task3_katelyn_toigo"
author: "Katelyn Toigo"
date: "10/31/2020"
output: html_document
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Summary

I will be analyzing the urchins.csv data.  I will be using version control, project management, data wrangling, analysis, and visualization skills. I will analyze urchin size vs. species (histogram, boxplot, and exploratory table) and counts by site and year (summary table and exploratory graph (histogram)).

## Read in the data

Read in the urchins.csv file and also clean the names so that they are all in lowercase_snakecase (to make working with the data easier).
```{r, echo=TRUE, message=FALSE, warning=FALSE}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Exploratory data visualization

### Faceted histogram

Comparing common_name vs. size.

Facet_wrap causes there to be two separate histograms, one for each common name.

labs() allows you to change the axis labels.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram() +
  facet_wrap(~common_name) +
  labs(x = "Size (cm)", y = "Count")
```

### Boxplot

Common name vs. size

labs() to label the axis

```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot()+
  labs(x = "Common Name", y = "Size (cm)")
```

## Descriptive Statistics Table

kable_styling() to make the graph format better in the knitted document
```{r, echo=TRUE, message=FALSE, warning=FALSE}
urchin_table <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n())
  kbl(urchin_table) %>% 
  kable_styling()
```
Purple urchins appear to be, on average, smaller in size than the red urchin (purple urchins have a mean that is 2.306692 cms smaller and their mean is 41% lower (this was calculated by finding the difference between the two means, dividing by their average and multiplying by 100)).  Purple urchins also have a smaller standard deviation and spread than the red urchins.  Both species have a relatively normal distribution when looking at the histogram.


## Explore purple urchin counts by site and year

Basic data wrangling - filter for only purple urchins, create a new columns called Date, year, and site_full.  Mutate is used to create the new columns.  mdy is used becasue the original data has dates in month, day, year order.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(year = year(Date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```


## Summary table containing total number of purple urchins by site and year

Group_by() is used first and followed by count() which basically combines group_by, summarize, and n all together in one function.
```{r, echo=TRUE, message=FALSE, warning=FALSE}

purple_table <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  count(site_full, year)
purple_table

```

## Exploratory graph of counts of purple urchins recorded at each site over time

Using a histogram set to 25 bins because it allowed for nice viewing between the years.  Facet wrap is used to separate each location into its own histogram.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = purple_urchins, aes(x = year, color = site_full)) +
  geom_histogram(bins=25) +
  facet_wrap(~site_full)
```









