---
title: "a3_task4_katelyn_toigo"
author: "Katelyn Toigo"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Summary

I will be analyzing the urchins.csv data.  I will be comparing the mean size of purple urchins in CONTROL (no giant kelp removal) or ANNUAL (annual kelp removal) sections of Naples Reef in January 2010.

## Read in the data

Read in the urchins.csv file and also clean the names so that they are all in lowercase_snakecase (to make working with the data easier).
```{r, echo=TRUE, message=FALSE, warning=FALSE}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```


## Wrangle the data

Filter to include only Naples Reef for January 2010.  Convert Date with lubridate, use month() and year() to create new month and year columns.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
purple_napl <- urchins %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(year = year(Date)) %>% 
  mutate(month = month(Date)) %>% 
  filter(site == "NAPL", year == 2010, month == 1)
```


## Graphs for better understanding of data

### Histogram assessing size distribution for each treatment

Size vs. Annual/Control
facet_wrap() for treatment because it then created two histograms - one for each type of treatment
```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = purple_napl, aes(x = size)) +
  geom_histogram() +
  facet_wrap(~treatment) +
  labs(x = "Size (cm)", y = "Count")
```

### QQ Plot assessing size distribution for each treatment

```{r, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = purple_napl, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment) +
  labs(x = "Size (cm)", y = "Count")
```

Main takeaways:  The annual treatment appears to have more spread in sizes and a lower kurtosis than the control treatment size distribution, when looking at the histograms. When looking at the qq-plots, both treatments are fairly linear meaning they are both close to being normally distributed. It looks like the annual plot is slightly more linear - slightly closer to normal distribution than the control.



## Table of descriptive statistics

Mean, median, standard deviation, sample size
kable_styling helps when formatting the knitted document.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
purple_napl_table <- purple_napl %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n())
  kbl(purple_napl_table) %>% 
  kable_styling()
```


## Two sample t-test

 This test is to compare means between the two groups (treatments).
 
 
 
 
 
 
 
 
 
 
